The Fates of Merging Supermassive Black Holes and A Proposal for a New Class of X-Ray Sources

### 2019-11-09 paper.tex
Finished bringing paper into compliance with format for MNRAS.

\begin{abstract}
We perform N-body simulations on 11 galaxies extracted from a cosmological simulation of hierarchical structure formation with total mass in the range $10^{12} \msun < M < 10^{13} \msun$ from $4 \ge z \ge 0$.  After mergers, we track the dynamical evolution of the orbiting black holes (BHs) around their host central BHs.  Of the 86 orbiting BHs, we find that 36 merge with their host central BHs, 13 are ejected from their respective host galaxies, and 37 are still orbiting at the end of the simulations.  Across all galaxies, 33 BHs are kicked to a higher orbit after close interactions with the central BHs, after which only one of them merged with their hosts.  These orbiting BHs should be detectable by their anomalous (not LMXRB) spectra.  The x-ray luminosities of the orbiting massive BHs at z=0 are in the range $10^{25}-10^{43}$ $erg\ s^{-1}$, with an undetectable median value $10^{32}$ $erg\ s^{-1}$.  However, the most luminous $\sim$$5\%$ should be detectable.
\end{abstract}

@@

Here we look at the merger history of 11 exemplary galaxies across the galaxy mass spectrum extracted from a cosmological simulation of hierarchical structure formation. We investigate how, after merging with incoming galaxies, SMBHs sink into the cores of the hosts and interact with the resident black hole. We show that gravitational interactions of multiple SMBHs are most probable in high-mass galaxies with total mass $10^{12} \msun < M < 10^{13} \msun$. Galaxies with lower masses have too few mergers with SMBH hosting galaxies. Galaxies with higher masses are more extended, making dynamical friction processes less efficient and hence failing to drive SMBHs into the host galaxy core.

@@

\begin{table}
\begin{center}
\caption{Galaxy characteristics at z=0}
\begin{tabular} {|c | c | c| c|}
\hline
$M_{DM}$ & $M_{CBH}$ & $M_{*} (orig.)$ & $M_{*} (rescaled)$ \\
\hline
 \num{2.85e13} 	&	 \num{1.46e9}	&	\num{5.58e12} 	&	\num{3.30e11}  \\
 \num{2.48e13} 	&	 \num{2.40e9}	&	\num{4.80e12} 	&	\num{2.25e11}  \\
 \num{2.29e13} 	&	 \num{1.22e9}	&	\num{4.46e12} 	&	\num{2.68e11}  \\
 \num{1.18e13} 	&	 \num{3.62e9}	&	\num{2.32e12} 	&	\num{2.83e11}  \\
 \num{1.03e13} 	&	 \num{9.85e9}	&	\num{1.96e12} 	&	\num{1.62e11}  \\
 \num{1.00e13} 	&	 \num{9.32e9}	&	\num{1.95e12} 	&	\num{2.49e11}  \\
 \num{8.92e12} 	&	 \num{5.60e9}	&	\num{1.74e12} 	&	\num{3.20e11}  \\
 \num{8.56e12} 	&	 \num{6.86e9}	&	\num{1.62e12} 	&	\num{1.83e11}  \\
 \num{6.94e12} 	&	 \num{2.02e9}	&	\num{1.72e12} 	&	\num{1.51e11}  \\
 \num{6.29e12} 	&	 \num{1.22e9}	&	\num{1.23e12} 	&	\num{1.87e11}  \\
 \num{1.84e12} 	&	 \num{2.21e9}	&	\num{5.52e11} 	&	\num{6.42e10}  \\
\hline
\end{tabular}
\end{center}
\label{table:gal_char}
\end{table}

@@

\subsection{Galaxy Gas X-Ray Luminosity}
We compute the x-ray luminosity of the hot gas in the galaxies, adapting Equation 26 from \citet{2012ApJ...754..125C}. In lieu of summing over discrete particles, we use the beta model of the gas as defined in our equation \ref{beta_model} and integrate:
\begin{equation}
\begin{aligned}
    L_x ={} \frac{\num{1.2e-24}}{(\mu m_p)^2}\left(\frac{kT_x}{1keV}\right)^{1/2}\rho_0^{2}\;4{\pi} \\
    \int_{0}^{\infty}\frac{r^2}{(1+r^2/r_c^2)^{3}}\;dr\text{  erg s}^{-1}.
\end{aligned}
\end{equation}

@@

The Eddington limit is defined as:
\begin{equation}
    L_{edd} = \num{1.26e38} (M/M_{\odot}) \text{  erg s}^{-1}.
\end{equation}

@@

\section{Conclusions}\label{sec:conclusions}
We performed N-body simulations on 11 galaxies ranging in mass between $10^{12} \msun < M < 10^{13} \msun$, from $4 \ge z \ge 0$.  We found that $42\%$ of the orbiting black holes merged with their host galaxy, while another $43\%$ remained in orbit at $z=0$.  With an x-ray luminosity range of $10^{25}-10^{43}$ $erg\ s^{-1}$

@@

\bibliographystyle{apj}
\bibliography{biblio}

%\begin{thebibliography}{}

%\bibitem[\protect\citeauthoryear{Aarseth}{2003}]{Aarseth03}
%Aarseth, S.~J., 2003, Gravitational N-Body Simulations (Cambridge University Press)

%\bibitem[\protect\citeauthoryear{K{\"u}pper et al.}{2011}]{Kupper11} 
%K{\"u}pper, A.~H.~W., Maschberger, T., Kroupa, P., Baumgardt, H., 2011, MNRAS, 417, 2300

%\end{thebibliography}
\end{document}

>>>>

\begin{abstract}
We perform N-body simulations on 11 galaxies extracted from a cosmological simulation of hierarchical structure formation with total mass in the range $10^{12} M_{\sun} < M < 10^{13} M_{\sun}$ from 4 $\gid$ z $\gid$ 0.  After mergers, we track the dynamical evolution of the orbiting black holes (BHs) around their host central BHs.  Of the 86 orbiting BHs, we find that 36 merge with their host central BHs, 13 are ejected from their respective host galaxies, and 37 are still orbiting at the end of the simulations.  Across all galaxies, 33 BHs are kicked to a higher orbit after close interactions with the central BHs, after which only one of them merged with their hosts.  These orbiting BHs should be detectable by their anomalous (not LMXRB) spectra.  The x-ray luminosities of the orbiting massive BHs at z=0 are in the range $10^{25}-10^{43}$ $erg\ s^{-1}$, with an undetectable median value $10^{32}$ $erg\ s^{-1}$.  However, the most luminous $\sim$$5\%$ should be detectable.
\end{abstract}

@@

Here we look at the merger history of 11 exemplary galaxies across the galaxy mass spectrum extracted from a cosmological simulation of hierarchical structure formation. We investigate how, after merging with incoming galaxies, SMBHs sink into the cores of the hosts and interact with the resident black hole. We show that gravitational interactions of multiple SMBHs are most probable in high-mass galaxies with total mass $10^{12} M_{\sun} < M < 10^{13} M_{\sun}$. Galaxies with lower masses have too few mergers with SMBH hosting galaxies. Galaxies with higher masses are more extended, making dynamical friction processes less efficient and hence failing to drive SMBHs into the host galaxy core.

@@

\begin{table}
\begin{center}
\caption{Galaxy characteristics at z=0. Units are M$\sun$.}
\begin{tabular} {|c | c | c| c|}
\hline
$M_{DM}$ & $M_{CBH}$ & $M_{*} (orig.)$ & $M_{*} (rescaled)$ \\
\hline
 2.85 $\times 10^{13}$ 	&	 1.46 $\times 10^9$	&	5.58 $\times 10^{12}$ 	&	3.30 $\times 10^{11}$  \\
 2.48 $\times 10^{13}$ 	&	 2.40 $\times 10^9$	&	4.80 $\times 10^{12}$ 	&	2.25 $\times 10^{11}$  \\
 2.29 $\times 10^{13}$ 	&	 1.22 $\times 10^9$	&	4.46 $\times 10^{12}$ 	&	2.68 $\times 10^{11}$  \\
 1.18 $\times 10^{13}$ 	&	 3.62 $\times 10^9$	&	2.32 $\times 10^{12}$ 	&	2.83 $\times 10^{11}$  \\
 1.03 $\times 10^{13}$ 	&	 9.85 $\times 10^9$	&	1.96 $\times 10^{12}$ 	&	1.62 $\times 10^{11}$  \\
 1.00 $\times 10^{13}$ 	&	 9.32 $\times 10^9$	&	1.95 $\times 10^{12}$ 	&	2.49 $\times 10^{11}$  \\
 8.92 $\times 10^{12}$ 	&	 5.60 $\times 10^9$	&	1.74 $\times 10^{12}$ 	&	3.20 $\times 10^{11}$  \\
 8.56 $\times 10^{12}$ 	&	 6.86 $\times 10^9$	&	1.62 $\times 10^{12}$ 	&	1.83 $\times 10^{11}$  \\
 6.94 $\times 10^{12}$ 	&	 2.02 $\times 10^9$	&	1.72 $\times 10^{12}$ 	&	1.51 $\times 10^{11}$  \\
 6.29 $\times 10^{12}$ 	&	 1.22 $\times 10^9$	&	1.23 $\times 10^{12}$ 	&	1.87 $\times 10^{11}$  \\
 1.84 $\times 10^{12}$ 	&	 2.21 $\times 10^9$	&	5.52 $\times 10^{11}$ 	&	6.42 $\times 10^{10}$  \\
\hline
\end{tabular}
\end{center}
\label{table:gal_char}
\end{table}

@@

\subsection{Galaxy Gas X-Ray Luminosity}
We compute the x-ray luminosity of the hot gas in the galaxies, adapting Equation 26 from \citet{2012ApJ...754..125C}. In lieu of summing over discrete particles, we use the beta model of the gas as defined in our equation \ref{beta_model} and integrate:
\begin{equation}
\begin{aligned}
    L_x ={} \frac{1.2 \times 10^{-24}}{(\mu m_p)^2}\left(\frac{kT_x}{1keV}\right)^{1/2}\rho_0^{2}\;4{\pi} \\
    \int_{0}^{\infty}\frac{r^2}{(1+r^2/r_c^2)^{3}}\;dr\text{  erg s}^{-1}.
\end{aligned}
\end{equation}

@@

The Eddington limit is defined as:
\begin{equation}
    L_{edd} = 1.26 \times 10^{38} (M/M_{\odot}) \text{  erg s}^{-1}.
\end{equation}

@@

\section{Conclusions}\label{sec:conclusions}
We performed N-body simulations on 11 galaxies ranging in mass between $10^{12}$ M$_{\sun}$ < M < $10^{13}$ M$_{\sun}$, from 4 $\gid$ z $\gid$ 0.  We found that 42$\%$ of the orbiting black holes merged with their host galaxy, while another 43$\%$ remained in orbit at $z=0$.  With an x-ray luminosity range of $10^{25}-10^{43}$ $erg\ s^{-1}$.

@@

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{biblio}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_guide.tex
